---
title: "AssignmentII"
author: "Group2"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction


# Data Exploration

# Question 1

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 2

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 3

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 4

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 5

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 6

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 7

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 8

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 9

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 10

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 11

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 12

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 

# Question 13

## Overview : 

## Code :

```{r echo=FALSE}

```

## Output : 

## Analysis : 
